---
title: "Tidy Tuesday 07-20-21"
author: "Geoffrey House"
date: "7/24/2021"
output:
  html_document:
    toc: true
    toc_depth: 3
---

### Lessons learned in working with this dataset:
* Using `lead` (and `lag` in `dplyr`) to calculate values in the current row using values either in subsequent (or previous) rows. Works in combination with `mutate`, `case_when` and `%in%`

#### The cumulative data correction is inconsistent among state/time entries, so using the categorical data instead (doesn't need the cumulative data correction)

```{r clean raw data, echo = FALSE, message = FALSE}
# Get the Data

rm(list = ls())

library(tidyverse)
library(janitor)

read_and_prep <- function(file, type){
  read_csv(paste0("./", file)) %>% 
    pivot_longer(cols = None:D4, names_to = "drought_lvl", values_to = type) %>% 
    janitor::clean_names()
  
}

state_area_pct <- read_and_prep("drought_percArea_2001-2021.csv", "area_pct")

# This fails
#state_area <- read_and_prep("drought_totArea_2001-2021.csv", "area_total")

state_pct_pop <- read_and_prep("drought_percPop_2001-2021.csv", "pop_pct")

# This fails
#state_pop <- read_and_prep("drought_totPop_2001-2021.csv", "pop_total")

all_df <- state_area_pct %>% 
  #left_join(state_area) %>% 
  left_join(state_pct_pop) %>% 
  #left_join(state_pop) %>% 
  rename(state_abb = state_abbreviation, stat_fmt = statistic_format_id)

all_df %>% 
  write_csv("./tabulated_categoricalDroughtData_2001_2021.csv")


# Get the Data

# Read in with tidytuesdayR package 
# Install from CRAN via: install.packages("tidytuesdayR")
# This loads the readme and all the datasets for the week of interest

# Either ISO-8601 date or year/week works!

#tuesdata <- tidytuesdayR::tt_load('2021-07-20')
#tuesdata <- tidytuesdayR::tt_load(2021, week = 30)

#drought <- tuesdata$drought

# Or read in the data manually

#drought <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-20/drought.csv')

#saveRDS(object = drought, file = "TidyTuesday_DroughtMapData.rds")

#drought <- readRDS(file = "TidyTuesday_DroughtMapData.rds")

```

```{r read in data, echo = F, message= F}

droughtData <- read_csv("./tabulated_categoricalDroughtData_2001_2021.csv")

```


#### The values for population and area affected by drought are cumulative across the 5 drought levels (D0-D4), i.e. values in D0 are totals of sum(D0:D4). To revert these to actual values, group by state and date, then subtract the next higher drought level data from each using the `lead` verb in `dplyr`

#### This wasn't working well (the cumulative calculation seemed inconsistent), so manually downloading the categorical (non-cumulative) data instead

```{r remove cumulative totals, echo = FALSE, message = FALSE}

# drought_actual <- drought %>% group_by(state_abb, map_date) %>%
#     mutate(actAreaPct = case_when(drought_lvl %in% c("None", "D4") ~ area_pct,
#                                   drought_lvl %in% c("D0","D1", "D2", "D3") ~ area_pct - lead(x = area_pct, n = 1)),
#            actAreaTot = case_when(drought_lvl %in% c("None", "D4") ~ area_total,
#                                   drought_lvl %in% c("D0","D1", "D2", "D3") ~ area_total - lead(x = area_total, n = 1)),
#            actPopPct = case_when(drought_lvl %in% c("None", "D4") ~ pop_pct,
#                                   drought_lvl %in% c("D0","D1", "D2", "D3") ~ pop_pct - lead(x = pop_pct, n = 1)),
#            actPopTot = case_when(drought_lvl %in% c("None", "D4") ~ pop_total,
#                                   drought_lvl %in% c("D0","D1", "D2", "D3") ~ pop_total - lead(x = pop_total, n = 1)))
# 


```

### Who is captured the most, and who has caught the villain the most?

* All characters have been captured between 70-90 times
* Daphnie and Velma have caught the villain far less than either Scooby (~4x less) or Fred (~x3 less)

```{r convert true false to logical, echo = F, message = F}

library(tidyr)
library(dplyr)
library(ggplot2)

namesToConvert <- names(scoobydoo)[18:32]

convertedData <- sapply(X = namesToConvert, FUN = function(name){return(as.logical(scoobydoo[[name]]))})

# Take the column sums to give the number of times each character was captured or caught the culprit
numberHolder <- data.frame(colSums(convertedData, na.rm = T))

# Need to transform to long format in order to plot. Split the row names on '_' to 
# get 1) whether the character was captured or caught the culprit, and 2) who the character is
actionHolder <- matrix(unlist(strsplit(rownames(numberHolder), split = "_")), ncol = 2, byrow = T)

numberHolder$action <- actionHolder[,1]
numberHolder$name <- actionHolder[,2]

names(numberHolder) <- c("numTimes", "action", "name")

# Remove the 'unmask' entries
holderForPlot <- numberHolder %>% filter(action != "unmask")

ggplot(data = holderForPlot, mapping = aes(x = numTimes, y = name)) + 
    geom_line(aes(group = name)) + 
    geom_point(aes(color = action), size =4) + 
    theme_bw()

```
