---
title: "Tidy Tuesday 07-20-21"
author: "Geoffrey House"
date: "7/24/2021"
output:
  html_document:
    toc: true
    toc_depth: 3
---

### Lessons learned in working with this dataset:
* Using `lead` (and `lag` in `dplyr`) to calculate values in the current row using values either in subsequent (or previous) rows. Works in combination with `mutate`, `case_when` and `%in%`
* How to embed html links in knit Rmd files

```{r clean raw data, echo = FALSE, message = FALSE}
# Get the Data

rm(list = ls())

library(tidyverse)
library(janitor)

read_and_prep <- function(file, type){
  read_csv(paste0("./", file)) %>% 
    pivot_longer(cols = None:D4, names_to = "drought_lvl", values_to = type) %>% 
    janitor::clean_names()
  
}

state_area_pct <- read_and_prep("drought_percArea_2001-2021.csv", "area_pct")

# This fails
#state_area <- read_and_prep("drought_totArea_2001-2021.csv", "area_total")

state_pct_pop <- read_and_prep("drought_percPop_2001-2021.csv", "pop_pct")

# This fails
#state_pop <- read_and_prep("drought_totPop_2001-2021.csv", "pop_total")

all_df <- state_area_pct %>% 
  #left_join(state_area) %>% 
  left_join(state_pct_pop) %>% 
  #left_join(state_pop) %>% 
  rename(state_abb = state_abbreviation, stat_fmt = statistic_format_id)

all_df %>% 
  write_csv("./tabulated_categoricalDroughtData_2001_2021.csv")


# Get the Data

# Read in with tidytuesdayR package 
# Install from CRAN via: install.packages("tidytuesdayR")
# This loads the readme and all the datasets for the week of interest

# Either ISO-8601 date or year/week works!

#tuesdata <- tidytuesdayR::tt_load('2021-07-20')
#tuesdata <- tidytuesdayR::tt_load(2021, week = 30)

#drought <- tuesdata$drought

# Or read in the data manually

#drought <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-20/drought.csv')

#saveRDS(object = drought, file = "TidyTuesday_DroughtMapData.rds")

#drought <- readRDS(file = "TidyTuesday_DroughtMapData.rds")

```



```{r read in data, echo = F, message= F}

droughtData <- read_csv(file = "./tabulated_categoricalDroughtData_2001_2021.csv")

```

```{r calculate changes in percentages between measurements, echo = F, message = F, warning = FALSE}

droughtData$drought_lvl <- as.factor(droughtData$drought_lvl)

# Need the as.numeric(NA) here because otherwise NA evaluates as a logical, not numeric, which is what's required for the other entries in the vector of the case_when.
droughtData_changes <- droughtData %>% group_by(state_abb, drought_lvl) %>%
    mutate(changeAreaPct = case_when(drought_lvl == "None" ~ as.numeric(NA),
                                     drought_lvl %in% c("D0", "D1", "D2", "D3","D4") ~ area_pct - lag(x = area_pct, n = 1, order_by = map_date)),
           changePopPct = case_when(drought_lvl == "None" ~ as.numeric(NA),
                                    drought_lvl %in% c("D0", "D1", "D2", "D3", "D4") ~ pop_pct - lag(x = pop_pct, n = 1, order_by = map_date)))


```

### With the current drought, more of the country is designated as under D4 (exceptional) than any time since the drought of 2012, [which was in turn the most widespread drought since 1956](https://www.ncdc.noaa.gov/sotc/drought/201206)

```{r graph changes over time by drought level area, echo = FALSE, message = FALSE, warning= FALSE}

library(ggplot2)
library(lubridate)

droughtData_changes %>% filter(drought_lvl != "None") %>% group_by(state_abb) %>%
    ggplot(data = ., mapping = aes(x = valid_start, y = changeAreaPct, group = state_abb)) + 
    geom_line(color = "#AAAAAA66") + geom_vline(xintercept = seq.Date(from = mdy("01-01-2000"), to = mdy("01-01-2022"), by = "years"), color = "#0000FF99") + theme_classic() + ggtitle("Drought severity for all states (1 gray line per state)") + xlab("Year") + ylab("Change in percentage of area \nwith this drought severity") + theme(axis.text.x=element_text(angle = -90, vjust = 0.5)) + facet_wrap(~drought_lvl, ncol = 1)

```

### There are very similar trends with the percentage of the population being affected by drought

```{r graph changes over time by drought level population, echo = FALSE, message = FALSE, warning= FALSE}

droughtData_changes %>% filter(drought_lvl != "None") %>% group_by(state_abb) %>%
    ggplot(data = ., mapping = aes(x = valid_start, y = changePopPct, group = state_abb)) + 
    geom_line(color = "#AAAAAA66") + geom_vline(xintercept = seq.Date(from = mdy("01-01-2000"), to = mdy("01-01-2022"), by = "years"), color = "#0000FF99") + theme_classic() + ggtitle("Drought severity for all states (1 gray line per state)") + xlab("Year") + ylab("Change in percentage of population affected \nwith this drought severity") + theme(axis.text.x=element_text(angle = -90, vjust = 0.5)) + facet_wrap(~drought_lvl, ncol = 1)

```

