---
title: "Tidy Tuesday 07-13-21"
author: "Geoffrey House"
date: "7/18/2021"
output:
  html_document:
    toc: true
    toc_depth: 3
---

### Lessons learned in working with this dataset:
* How to make an easy Cleveland dot plot in ggplot using data groupings

```{r get data, echo = FALSE, message = FALSE}
# Get the Data

rm(list = ls())

# Read in with tidytuesdayR package 
# Install from CRAN via: install.packages("tidytuesdayR")
# This loads the readme and all the datasets for the week of interest

# Either ISO-8601 date or year/week works!

#tuesdata <- tidytuesdayR::tt_load('2021-07-13')
#tuesdata <- tidytuesdayR::tt_load(2021, week = 29)

#scoobydoo <- tuesdata$scoobydoo

# Or read in the data manually

#scoobydoo <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-13/scoobydoo.csv')

#saveRDS(object = scoobydoo, file = "TidyTuesday_ScoobyDoo.rds")

scoobydoo <- readRDS(file = "TidyTuesday_ScoobyDoo.rds")

```

### Who is captured the most, and who has caught the villain the most?

* All characters have been captured between 70-90 times
* Daphnie and Velma have caught the villain far less than either Scooby (~4x less) or Fred (~x3 less)

```{r convert true false to logical, echo = F, message = F}

library(tidyr)
library(dplyr)
library(ggplot2)

namesToConvert <- names(scoobydoo)[18:32]

convertedData <- sapply(X = namesToConvert, FUN = function(name){return(as.logical(scoobydoo[[name]]))})

# Take the column sums to give the number of times each character was captured or caught the culprit
numberHolder <- data.frame(colSums(convertedData, na.rm = T))

# Need to transform to long format in order to plot. Split the row names on '_' to 
# get 1) whether the character was captured or caught the culprit, and 2) who the character is
actionHolder <- matrix(unlist(strsplit(rownames(numberHolder), split = "_")), ncol = 2, byrow = T)

numberHolder$action <- actionHolder[,1]
numberHolder$name <- actionHolder[,2]

names(numberHolder) <- c("numTimes", "action", "name")

# Remove the 'unmask' entries
holderForPlot <- numberHolder %>% filter(action != "unmask")

ggplot(data = holderForPlot, mapping = aes(x = numTimes, y = name)) + 
    geom_line(aes(group = name)) + 
    geom_point(aes(color = action), size =4) + 
    theme_bw()

```

### How have different exclamations changed over time?

* *'Just about wrapped up'*, *'Groovy'*, and *'Jeepers!'* were all barely used from 1980-2000, but are back now
* *'Zoinks!'* has always been the most popular, and *'Scooby Doo, where are you?'* isn't actually said all that often.

```{r correlations between words used and imdb ratings, echo = F, message = F, warning= F}

colsToConvertNumeric <- names(scoobydoo)[58:65]

convertedDataSayings <- as.data.frame(sapply(X = colsToConvertNumeric, FUN = function(entry){return(as.numeric(scoobydoo[[entry]]))}))

convertedDataSayings$date <- scoobydoo$date_aired

convertedDataSayings$year <- lubridate::year(convertedDataSayings$date)

convertedDataSayings_long <- tidyr::pivot_longer(data = convertedDataSayings, cols = c(-year, -date), names_to = "saying", values_to = "Number of times said")

# Convert 0 entries to NA
convertedDataSayings_long$`Number of times said`[which(convertedDataSayings_long$`Number of times said` == 0)] <- NA

ggplot(data = convertedDataSayings_long, mapping = aes(x = date, y = saying, size = `Number of times said`)) + geom_point(aes(color = saying)) + 
    theme_bw() + xlab("Year") + ylab("Saying") + scale_color_brewer(palette = "Set2", guide = "none")

```
