---
title: "Tidy Tuesday 06-29-21"
author: "Geoffrey House"
date: "7/3/2021"
output:
  html_document:
    toc: true
    toc_depth: 3
---

### Lessons learned in working with this dataset:
* In `ggplot`, the coordinate labels are also flipped after using `coord_flip()` (i.e. `xlab` specifies the y-axis label, and vice-versa)
* Looking up latitude/longitude entries using Geocoding from `tmaptools`
* Downloading base maps (and picking the right bounding box/zoom level) using ggmap
* Plotting locations on a map

```{r get data, echo = FALSE, message = FALSE}
# Get the Data

library(dplyr)
library(ggplot2)
library(tmaptools)
library(here)
library(tidyr)
library(ggmap)

rm(list = ls())

# Read in with tidytuesdayR package 
# Install from CRAN via: install.packages("tidytuesdayR")
# This loads the readme and all the datasets for the week of interest

# Either ISO-8601 date or year/week works!

#tuesdata <- tidytuesdayR::tt_load('2021-06-29')
# #tuesdata <- tidytuesdayR::tt_load(2021, week = 26)
# 
#animalRescues <- tuesdata$animal_rescues
# 
# saveRDS(object = animalRescues, file = "animalRescuesData_062921.rds")

animalRescues_orig <- readRDS(file = "animalRescuesData_062921.rds")

animalRescues <- animalRescues_orig

```

```{r clean data, echo = FALSE, message = FALSE}

# Make the animal labels consistent
animalRescues$animal_group_parent[which(animalRescues$animal_group_parent == "cat")] <- "Cat"


```

#### By far, most of the rescues are for cats, with birds as the second most rescued, and dogs the third.

```{r animal type histogram, echo = FALSE, message = FALSE}

animalRescues %>% ggplot(mapping = aes(x = animal_group_parent)) + geom_bar() + theme_bw() + 
    ylab("Number of rescues") + xlab("Type of animal rescued") + coord_flip()

```


```{r assign lat lon to boroughs, echo = FALSE, message = FALSE}

# Fully functional, but only needed once, then saved as RDS. Reads the RDS file at the end of the code block.

# For some rescues, there are specific lat/long values, but for others there aren't. Assign those missing
# lat/lon values to the be lat/lon of the borough in general

# First need to make all of the borough names lower case for consistency
# animalRescues$borough <- tolower(animalRescues$borough)
# 
# assignLatLon <- function(boroughQuery){
#     # Geocoding of lat/lon using openStreetMap
#     OSM_returned <- tmaptools::geocode_OSM(boroughQuery)
#     if(!is.null(OSM_returned)){
#         OSM_lat <- OSM_returned$coords[2]
#         OSM_lon <- OSM_returned$coords[1]
# 
#         dataReturn <- data.frame("lat" = OSM_lat, "lon" = OSM_lon)
#     } else{
#         print(paste0("Lookup failed for query: ", boroughQuery))
#         dataReturn <- data.frame("lat" = NA, "lon" = NA)
#     }
#     return(dataReturn)
# }
# 
# uniqueBoroughs <- data.frame("boroughName" = unique(animalRescues$borough),
#                              "latitude" = vector(mode = "numeric", length = length(unique(animalRescues$borough))),
#                              "longitude" = vector(mode = "numeric", length = length(unique(animalRescues$borough))), stringsAsFactors = FALSE)
# 
# # Geocode the city names to convert into lat/lon
# for(rowNum in seq(1, nrow(uniqueBoroughs))){
#     #print(rowNum)
# 
#     currBorough <- uniqueBoroughs$boroughName[rowNum]
#     print(currBorough)
#     returnedGeocode <- assignLatLon(currBorough)
#     uniqueBoroughs$latitude[rowNum] <- returnedGeocode$lat
#     uniqueBoroughs$longitude[rowNum] <- returnedGeocode$lon
# 
#     # Not supposed to exceed 1 query per second for open streets mapping
#     Sys.sleep(2)
# 
# }
# 
# # Need to clean up the entries for sutton, camden, bexley, brent, and brentwood because the lookups didn't work right. fix by adding a temporary 'uk' addition to the name
# 
# fixIndex <- which(uniqueBoroughs$boroughName %in% c("sutton", "camden", "bexley", "brent", "brentwood"))
# 
# for(rowNum2 in fixIndex){
#     fixedName <- paste0(uniqueBoroughs$boroughName[rowNum2], ", uk")
#     fixedGeocode <- assignLatLon(fixedName)
#     uniqueBoroughs$latitude[rowNum2] <- fixedGeocode$lat
#     uniqueBoroughs$longitude[rowNum2] <- fixedGeocode$lon
# }
# 
# # now remove the NA row
# uniqueBoroughs <- uniqueBoroughs[which(!is.na(uniqueBoroughs$boroughName)),]
# 
# # Add the borough lat/lon for any rescue row missing a lat/lon
# for(rowNum3 in seq(1,nrow(animalRescues))){
#     currLat <- animalRescues$latitude[rowNum3]
#     currLon <- animalRescues$longitude[rowNum3]
#     #print(currLat)
#     #print(currLon)
#     if(currLat != "NULL" && currLat != "0" && currLon != "NULL" && currLon != "0"){
#         next
#      
#     } else{
#         #print("replace")
#         currBorough <- animalRescues$borough[rowNum3]
#         boroughLat <- uniqueBoroughs$latitude[which(uniqueBoroughs$boroughName == currBorough)]
#         boroughLon <- uniqueBoroughs$longitude[which(uniqueBoroughs$boroughName == currBorough)]
#         animalRescues$latitude[rowNum3] <- boroughLat
#         animalRescues$longitude[rowNum3] <- boroughLon
#     }
#     
# }
# 
# # Convert the lat/lon columns to numeric
# animalRescues$latitude <- as.numeric(animalRescues$latitude)
# animalRescues$longitude <- as.numeric(animalRescues$longitude)
# 
# animalRescues_locsAdded <- animalRescues
# 
# saveRDS(object = animalRescues_locsAdded, file = "animalRescuesData_locsAdded_062921.rds")

animalRescues_locsAdded <- readRDS(file = "animalRescuesData_locsAdded_062921.rds")
```

```{r prep maps for rescues, echo = FALSE, message = FALSE}

# Functional, just commented out except the last line reading the base maps from the RDS

# hist(animalRescues$latitude)
# 
# # Add room around bounding box for map coordinates
# 
# minLat <- 51
# maxLat <- 52
# minLon <- -1
# maxLon <- 1
# 
# # minLat <- round(x = min(animalRescues_locsAdded$latitude, na.rm = T)) - 1
# # maxLat <- round(max(animalRescues_locsAdded$latitude, na.rm = T)) + 1
# # minLon <- round(min(animalRescues_locsAdded$longitude, na.rm = T)) - 1
# # maxLon <- round(max(animalRescues_locsAdded$longitude, na.rm = T)) + 1
# 
# boundBox <- c(minLon, minLat, maxLon, maxLat)
# 
# # # Toner with no labels, but does have state boundaries
# london_map_toner_background <- ggmap::get_stamenmap(bbox = boundBox, crop = T, zoom = 10, color = "color", maptype = "toner-background")
# # Watercolor with no state boundaries
# london_map_watercolor <- ggmap::get_stamenmap(bbox = boundBox, crop = T, zoom = 10, color = "color", maptype = "watercolor")
# 
# # 
# saveRDS(object = london_map_toner_background, file = "London_backgroundMap_stamenTonerBackground.Rds")
# saveRDS(object = london_map_watercolor, file = "London_backgroundMap_stamenWatercolor.Rds")

london_map_toner_background <- readRDS(file = "London_backgroundMap_stamenTonerBackground.Rds")
london_map_watercolor <- readRDS(file = "London_backgroundMap_stamenWatercolor.Rds")

```

#### Cats need rescuing all over the place!

```{r plot map cat rescue, echo = FALSE, message = FALSE}

# Get only the cat rescue entries
catRescues <- animalRescues_locsAdded %>% filter(animal_group_parent == "Cat") %>% group_by(latitude, longitude) %>% summarise(`Number of rescues` = n())

# 35% transparency
ggmap(london_map_toner_background) + geom_point(data = catRescues, mapping = aes(x = longitude, y = latitude, size = `Number of rescues`), color = "#c51b7d59") + xlab("longitude") + ylab("latitude") + ggtitle("Cat rescues in greater London")

```

#### Most bird rescues seem centered near the center of London (darker purple)

```{r plot map bird rescue, echo = FALSE, message = FALSE}

# Get only the bird rescue entries
birdRescues <- animalRescues_locsAdded %>% filter(animal_group_parent == "Bird") %>% group_by(latitude, longitude) %>% summarise(`Number of rescues` = n())

# 35% transparency
ggmap(london_map_toner_background) + geom_point(data = birdRescues, mapping = aes(x = longitude, y = latitude, size = `Number of rescues`), color = "#762a8359") + xlab("longitude") + ylab("latitude") + ggtitle("Bird rescues in greater London")

```

#### There are far fewer dog rescues, but they're more geographically widespread like the cat rescues

```{r plot map dog rescue, echo = FALSE, message = FALSE}

# Get only the dog rescue entries
dogRescues <- animalRescues_locsAdded %>% filter(animal_group_parent == "Dog") %>% group_by(latitude, longitude) %>% summarise(`Number of rescues` = n())

# 35% transparency
ggmap(london_map_toner_background) + geom_point(data = dogRescues, mapping = aes(x = longitude, y = latitude, size = `Number of rescues`), color = "#4d922159") + xlab("longitude") + ylab("latitude") + ggtitle("Dog rescues in greater London")

```

#### There seem to be a lot of foxes living (and rescued!) even in the center of London. (The larger dots represent rescues where the location wasn't exactly known, and was guessed)

```{r plot map fox rescue, echo = FALSE, message = FALSE}

# Get only the fox rescue entries
foxRescues <- animalRescues_locsAdded %>% filter(animal_group_parent == "Fox") %>% group_by(latitude, longitude) %>% summarise(`Number of rescues` = n())

# 35% transparency
ggmap(london_map_toner_background) + geom_point(data = foxRescues, mapping = aes(x = longitude, y = latitude, size = `Number of rescues`), color = "#b3580659") + xlab("longitude") + ylab("latitude") + ggtitle("Fox rescues in greater London")

percNullLocs_fox <- animalRescues_orig %>% filter(animal_group_parent == "Fox") %>% mutate(totalNum = n()) %>% filter(latitude == "NULL") %>% mutate(numNull = n()) %>% select(totalNum, numNull) %>% unique() %>% summarise(percNullLocs_fox = (numNull/totalNum) * 100)

percNullLocs_cat <- animalRescues_orig %>% filter(animal_group_parent == "Cat") %>% mutate(totalNum = n()) %>% filter(latitude == "NULL") %>% mutate(numNull = n()) %>% select(totalNum, numNull) %>% unique() %>% summarise(percNullLocs_cat = (numNull/totalNum) * 100)

```